A Fast and Tight Heuristic for A* in Road Networks

### 2021-01-17 paper/ch_potentials.tex
readding turn costs
Consistent heuristic
formal setup

Unfortunately, for many real world applications, this basic graph model is too simplistic.
Many additional features need to be considered.
% --- ich habe turn costs weg gelassen, weil wir die nicht in den experimenten haben, ich will keiner Reviewer auf blöde Ideeen bringen
% --- ich habe temporary driving bans weg gelassen, weil wir das nicht in den Experimenten haben und ich "we can do all of the above features" claimen möchte
This includes turn restrictions, live traffic, user preferences, traffic predictions.%, temporary driving bans for certain types of vehicles.
Specific applications often have additional application-specific requirements.
Extending Dijkstra's algorithm to support these features is usually easy.
On the other hand, while extending hierarchical speed-up techniques is possible, the algorithm development is vastly more complex.
For every feature, dedicated research paper(s) exist that extend CH.
Supporting the combination of several features is even harder. 
For example, we are not aware of any work combining all features mentioned above.
In this paper, we describe an algorithm that handles the combination of all mentioned features.

@@

@@

The query phase then is to execute A* on $G_q$.
Before running the search, the heuristic will be initialized with the target $T(t)$.
The distance estimate for a node $x$ from $G_q$ can be computed as $h_{T(t)}(T(x))$.
For CH-Potentials to find correct distances, $w_q \geq w_\ell$ has to hold for all arcs.
\todo{Correctness Argument?}
\todo{Describe A*?}

>>>>

Unfortunately, for many real world applications, this basic graph model is too simplistic.
Many additional features need to be considered.
% --- ich habe temporary driving bans weg gelassen, weil wir das nicht in den Experimenten haben und ich "we can do all of the above features" claimen möchte
This includes turn costs and restrictions, live traffic, user preferences, traffic predictions.%, temporary driving bans for certain types of vehicles.
Specific applications often have additional application-specific requirements.
Extending Dijkstra's algorithm to support these features is usually easy.
On the other hand, while extending hierarchical speed-up techniques is possible, the algorithm development is vastly more complex.
For every feature, dedicated research paper(s) exist that extend CH.
Supporting the combination of several features is even harder. 
For example, we are not aware of any work combining all features mentioned above.
In this paper, we describe an algorithm that handles the combination of all mentioned features.

@@

\subsection{Formal Setup: Inputs, Outputs, and Phases}

Throughout this paper, we consider a two phase setup with an off-line preprocessing phase and an on-line query phase.
The input to the preprocessing phase is data that is static, such as the road network and freeflow travel times.
On the other hand, the input to the query phase changes a lot.
For example, the source and the target nodes are query inputs.
Further, the user preferences and live traffic are query inputs.
The preprocessing phase may be slow and computes auxiliary data.
The query phase obtains this auxiliary data as input and performs the path computation.

Formally, the input to the preprocessing phase is a graph $G_\ell$ with lower bound weights $w_\ell$.
$w_\ell(e)$ must be a scalar value for every edge $e$ of $G_\ell$.
In this paper, $w_\ell$ is always the freeflow travel time.
The output is auxiliary data that enables an efficient heuristic function $h_t(x)$.
$h_t(x)$ is the exact distance from $x$ to $t$ in $G_\ell$.

The formal inputs of the query phase depend on the considered features.
In the simplest case, the input is a query weight $w_q$, and source and target nodes $s$ and $t$.
It is guaranteed that $w_\ell(e) \le w_q(e)$ for all edges $e$.
The output is a shortest $st$-path with respect to $w_q$.
The query has access to the $h_t$ heuristic function.
An example of the simple case is live-traffic.
In this case, $w_q$ is an array of live travel times.

With more complex features, the inputs become more complex.
The extended query weight $w_q(e,\tau,p)$ has three parameters: the edge $e$, the edge entry time $\tau$, and the user preferences $p$.
Beside $s$ and $t$ the input contains a departure time $\tau_s$.
The distance of a path $e_1,e_2\ldots$ is defined as $w_q(e_1,\tau_s,p)+w_q(e_2,\tau_s+w_q(e_1,\tau_s,p),p)+\cdots$
Further, the query input contains a graph $G_q$ with nodes and edges differing from $G_\ell$.
To be able to use the $h$ heuristic, the query implementation need to have two node mapping functions $\phi$ and $\psi$ that relate nodes in $G_\ell$ and $G_q$ as follows:
$h_{\phi(t)}(\psi(x))$ must lower bound the distance from $x$ to $t$ in $G_q$.

Applications requiring many features have an inherent complexity.
The departure and entry times are needed for prediction traffic.
The user preference parameter is needed for tunnel and highway avoidance.
The differentiation between $G_\ell$ and $G_q$ is needed to support turn costs and restrictions.

This application complexity translates into the complex query inputs. 
Fortunately, the preprocessing inputs are unaffected.
A highlight of our technique is that all features can be supported without changing the preprocessing at all.

@@

The query phase then is to execute A* on $G_q$.
Before running the search, the heuristic will be initialized with the target $T(t)$.
The distance estimate for a node $x$ from $G_q$ can be computed as $h_{T(t)}(T(x))$.
For CH-Potentials to find correct distances, $w_q \geq w_\ell$ has to hold for all arcs.
\todo{Correctness Argument? Stichwort: Consistent/Montone heuristic}
\todo{Describe A*?}
